European or American? A re-examination of the transatlantic influence over the IMF

### 2025-05-22 manuscript/manuscript.qmd
extra lit review
American influence
describe PCA sources

>>>>

## Independent Variables

To assess the influence of the United States and European powers on IMF lending decisions, we construct two key independent variables that capture each actor's *propensity* to exert influence in favour of a country's borrowing from the IMF at a given time. These propensities are based on widely used proxies from the existing literature, which reflect the degree of a country’s political and economic alignment with the US or Europe [@lipscy2018]. This approach assumes that both the US and European powers are more likely to advocate for lending decisions that benefit their close partners.

Specifically, we construct our independent variables using three indicators commonly interpreted as channels of political and economic affinity, similar to @lipscy2018:

-   UN General Assembly voting similarity[^3]
-   Bilateral trade intensity[^4]
-   Banking sector exposure[^5]

[^3]: Originallly used in @thacker1999.

[^4]: From @barro2005.

[^5]: See @broz2006; @lipscy2017; @oatley2004.
